---
title: "DataAnalysis - Postprocessing Results MaxQuant"
autor: David Lilek
date: "`r format(Sys.time(), '%d %B %Y, %X')`"
output:
  html_document:
    df_print: paged
    fig.align: center
    self_contained: yes
    fig.height: 4
    fig.width: 8
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

# Global Parameters 

```{r}
###############################
#
#define settings
#
###############################

GET_SAMPLE_NAMES <- FALSE # if set to FALSE define it in the next line
sample_names_user <- sample(LETTERS,1)
unique <- 2  #how many peptides should be at least necessary to identify a protein

RAZOR <- TRUE #TRUE | FALSE
UNIQUE <- TRUE #TRUE | FALSE
LFQ <- TRUE #TRUE | FALSE

path = "N:/1_A_Bachelor_Master_Intern/00_M_2022/David/Data/3_20220406_TR/20220406_TR_QC_MBR_pooled_rerun_4masterthesis/"
#path = "C:/Users/lilek/Desktop/20220406_TR_QC_MBR_rerun_4masterthesis/"

```


# Introduction

## LFQ-Procedure

Filtering:

-   `Potential.contaminant == ""` -> remove where contaminants
-   `Reverse == ""`-> remove where reverse = +
-   `Only.identified.by.site == ""` -> remove where only identified by site = +

Then the `log2`of the LFQ intensities is calculated and all `-Inf` values were substituted by NA using the following command: `lfq[lfq == -Inf] <- NA`

Then the number of lfq values is used for the further evaluation, visualization and statistical testing. In the following this data will be called LFQ.

## Unique | Razor&Unique procedure

Filtering:

-   `Potential.contaminant == ""` -> remove where contaminants
-   `Reverse == ""`-> remove where reverse = +
-   `Only.identified.by.site == ""` -> remove where only identified by site = +

-   `data[name] >= unique` -> keep only identifications which have greater or equal number (defined via variable `unique`) of `razor/unique peptides ` | typically this is set to 1 or 2


# Load required packages and functions

```{r start, message=FALSE, warning=FALSE, include=FALSE}
start_time = format(Sys.time(), '%X')
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
#load libraries
library(tidyverse)
library(dplyr)
library(gplots)
library(reshape2) 
library(pheatmap)
#load functions
filtering <- function(data_raw){
  data_raw %>% filter( 
    Potential.contaminant == "",
    Reverse == "",
    Only.identified.by.site == "")
}

CV <- function(x){
  (sd(x)/mean(x))*100
}
```

# Data Analysis

The txt-files are read in, processed and filtered automatically. This section shows the code, heatmaps for LFQ values and also some summary statistics which are only shown for further discussion on the results, for the documentation of the statistical evaluation and are not labeled/discussed yet. 

```{r, results="asis"}
##########read in all data with pattern .txt
# get file list
file.list <- list.files(path = path, pattern='*.txt')
# create file path
file.path <- paste(path,file.list,sep = "")
# read in files - for each file one variable is created - results are stored in a list
res_raw <- lapply(file.path, read.delim)
names(res_raw) <- file.list

if (GET_SAMPLE_NAMES == TRUE){
  #get sample names
  tmp <- colnames(res_raw[[1]])
  tmp <- tmp[grep("Razor...unique.peptides.",tmp)]
  tmp <- sub(".*Razor...unique.peptides.", "",tmp)
  sample_names_raw <- tmp
  rm(tmp)
  #create sample names
  sample_names_razor <- paste("Razor...unique.peptides.",sample_names_raw, sep="")
  sample_names_unique <- paste("Unique.peptides.",sample_names_raw, sep="")
  sample_names_lfq <- paste("LFQ.intensity.",sample_names_raw, sep="")
} else {
  for (i in 1:length(file.list)){
    #rename columns
    res_raw[[i]] <- res_raw[[i]] %>%
       rename_with(~ paste("Razor_", sample_names_user, sep = ""), starts_with("Razor...unique.peptides.")) %>%
       rename_with(~ paste("Unique_", sample_names_user, sep = ""), starts_with("Unique.peptides.")) %>%
       rename_with(~ paste("LFQ_", sample_names_user, sep = ""), starts_with("LFQ.intensity."))
    #get column names
    sample_names_razor <- colnames(res_raw[[i]][grep("Razor_", colnames(res_raw[[i]]))])
    sample_names_unique <- colnames(res_raw[[i]][grep("Unique_", colnames(res_raw[[i]]))])
    sample_names_lfq <- colnames(res_raw[[i]][grep("LFQ_", colnames(res_raw[[i]]))])
}
}

#another option would be
#uniq <- data %>% select(starts_with("Unique.peptides."))



#create data frame for razor
results_2gether_razor <- data.frame(matrix(ncol = length(sample_names_razor), nrow = length(file.list)))
row.names(results_2gether_razor) <- file.list
colnames(results_2gether_razor) <- sample_names_razor
#create data frame for unique
results_2gether_unique <- data.frame(matrix(ncol = length(sample_names_unique), nrow = length(file.list)))
row.names(results_2gether_unique) <- file.list
colnames(results_2gether_unique) <- sample_names_unique
#create data frame for lfq
results_2gether_lfq <- data.frame(matrix(ncol = length(sample_names_lfq), nrow = length(file.list)))
row.names(results_2gether_lfq) <- file.list
colnames(results_2gether_lfq) <- sample_names_lfq

#results lists
results_razor <- list()
results_unique <- list()
results_lfq <- list()

if (RAZOR == TRUE) {
  cat("\n")
  cat("##","Razor","\n")
  cat("\n")
for (x in file.list){
  cat("\n")
  cat("###",substring(x, 1, nchar(x)-4),"\n")
  data_raw <- as.data.frame(res_raw[[x]])
  data <- filtering(data_raw)
  results <- data.frame(matrix(ncol = length(sample_names_razor), nrow = nrow(data)))
  colnames(results) <- sample_names_razor
  i <- 0
  for (i in 1:length(sample_names_razor)){
    name <- sample_names_razor[i]
    tmp <- (data[name] >= unique)
    results[i] <- tmp
    cat(name,"<br/>",sum(tmp, na.rm = TRUE),"<br/>")
  }
  
  results_razor[[x]] <- results
  #assign(paste("results_razor",x,sep="_"),results) #create individual variable names

  results_2gether_razor[x,] <- sapply(results,function(x)sum(x, na.rm = TRUE))
  cat("\n")
}
}

if (UNIQUE == TRUE) {
  cat("\n")
  cat("##","Unique","\n")
  cat("\n")
for (x in file.list){
  cat("\n")
  cat("###",substring(x, 1, nchar(x)-4),"\n")
  data_raw <- as.data.frame(res_raw[[x]])
  data <- filtering(data_raw)
  results <- data.frame(matrix(ncol = length(sample_names_unique), nrow = nrow(data)))
  colnames(results) <- sample_names_unique
  i <- 0
  for (i in 1:length(sample_names_unique)){
    name <- sample_names_unique[i]
    tmp <- (data[name] >= unique)
    results[i] <- tmp
    cat(name,"<br/>",sum(tmp, na.rm = TRUE),"<br/>")
  }
  
  results_unique[[x]] <- results
  #assign(paste("results_razor",x,sep="_"),results) #create individual variable names

  results_2gether_unique[x,] <- sapply(results,function(x)sum(x, na.rm = TRUE))
  cat("\n")
}
}

if (LFQ == TRUE) {
    cat("\n")
  cat("##","LFQ","\n")
  cat("\n")
  for (x in file.list){
    cat("\n")
    cat("###",substring(x, 1, nchar(x)-4),"\n")
    data_raw <- as.data.frame(res_raw[[x]])
    data <- filtering(data_raw)
    
    results <- data.frame(matrix(ncol = length(sample_names_lfq), nrow = nrow(data)))
    colnames(results) <- sample_names_lfq
    i <- 0
    for (i in 1:length(sample_names_lfq)){
      name <- sample_names_lfq[i]
      tmp <- log2(data[name])
      tmp[tmp == -Inf] <- NA
      results[i] <- tmp
      cat(name,"<br/>",sum(!is.na(tmp)),"<br/>")
      
    }
    
    results_lfq[[x]] <- results
    #assign(paste("results_lfq",x,sep="_"),results) #create individual variable names
    #create heatmaps
    paletteLength <- 50
    myColor <- colorRampPalette(c("navy", "white",  "firebrick3"))(paletteLength)
      lfq4pheatmap <- results
      if (ncol(lfq4pheatmap) < 2){
        cat("\n")
        cat("##"," No heatmap possible. Too less columns in the result table.","\n")
      } else {
      lfq4pheatmap[is.na(lfq4pheatmap)] <- 0
      
      pheatmap(lfq4pheatmap,
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_row = rep("",nrow(lfq4pheatmap)))

      }
    results_2gether_lfq[x,] <- sapply(results,function(x)sum(!is.na(x)))
    cat("\n")
    
  }
}

#add number in column 4 for selection
results_2gether_razor$number <- c(1:length(file.list))
results_2gether_unique$number <- c(1:length(file.list))
results_2gether_lfq$number <- c(1:length(file.list))

#remove variables - clean up
rm(data,
   data_raw,
   name,
   results,
   sample_names_razor,
   sample_names_unique,
   sample_names_lfq,
   tmp,
   x)

```

# Used time

```{r end, message=FALSE, warning=FALSE, include=FALSE}
end_time = format(Sys.time(), '%X')
```

```{r used time, echo=FALSE, message=TRUE, warning=FALSE}
print(paste("Time used for analysis:", round(as.difftime(end_time, units = "mins")-as.difftime(start_time, units = "mins"),digits=2),"minutes"))
```
# Summary



```{r}
if (RAZOR == TRUE){
  results_2gether_razor
}
if (UNIQUE == TRUE){
  results_2gether_unique
}
if (LFQ == TRUE){
  results_2gether_lfq
}
```


# OLD STUFF



```{r eval=FALSE, include=FALSE}
razor_comp <- results_2gether_razor[c(1,3,15,25,27,29),-4]
print(razor_comp)
boxplot(t(razor_comp), 
        names=seq(1:6),
        xlab="Experiment No",
        ylab="Identifications Razor",
        cex.lab = 0.8)

res_new <- t(razor_comp[c(1:3),])
res_old <- t(razor_comp[c(4:6),])
#print(res_new-res_old)

shapiro.test(res_new-res_old)

t <- t.test(res_new-res_old,
       µ=0)
print(t)
```



```{r eval=FALSE, include=FALSE}
lfq_comp <- results_2gether_lfq[c(1,3,15,25,27,29),-4]
print(lfq_comp)
boxplot(t(lfq_comp), 
        names=seq(1:6),
        xlab="Experiment No",
        ylab="Identifications LFQ",
        cex.lab = 0.8)

res_new <- t(lfq_comp[c(1:3),])
res_old <- t(lfq_comp[c(4:6),])
#print(res_new-res_old)

shapiro.test(res_new-res_old)

t.test(res_new-res_old,
       µ=0)
```



```{r eval=FALSE, include=FALSE}
boxplot(results_2gether_razor[c(5:14),-4])
rsd_002 <- apply(results_2gether_razor[c(5:14),-4],1,CV)
rsd_003 <- apply(results_2gether_razor[c(15:24),-4],1,CV)

boxplot(rsd_002,rsd_003,
        xlab="rerun experiment no.",
        names=c("002","003"),
        ylab="relative sd")
```


.

```{r eval=FALSE, include=FALSE}
boxplot(results_2gether_lfq[c(5:14),-4])
boxplot(results_2gether_lfq[c(15:24),-4])
rsd_002 <- apply(results_2gether_lfq[c(5:14),-4],1,CV)
rsd_003 <- apply(results_2gether_lfq[c(15:24),-4],1,CV)

boxplot(rsd_002,rsd_003,
        xlab="rerun experiment no.",
        names=c("002","003"),
        ylab="relative sd")
```


```{r eval=FALSE, include=FALSE}
rsd_002_r <- apply(results_2gether_razor[c(5:14),-4],1,CV)
rsd_003_r <- apply(results_2gether_razor[c(15:24),-4],1,CV)
rsd_002 <- apply(results_2gether_lfq[c(5:14),-4],1,CV)
rsd_003 <- apply(results_2gether_lfq[c(15:24),-4],1,CV)
```


```{r eval=FALSE, include=FALSE}
razor <- results_2gether_razor[c(5:24),-4]
lfq <- results_2gether_lfq[c(5:24),-4]

factor <- factor(c(rep("002_razor",10),rep("003_razor",10)))
razor$factor <- factor
factor <- factor(c(rep("002_lfq",10),rep("003_lfq",10)))
lfq$factor <- factor

data_tmp <- rbind (razor,lfq)

dat.m <- melt(data_tmp, measure.vars=c('A1','A2','A3'))

ggplot(subset(dat.m, factor == "002_razor" | factor == "002_lfq")) +
  geom_boxplot(aes(x=variable, y=value, color=factor)) +
  labs(title="Comparison Rerun LFQ-Razor Experiment 002",
       x ="Extract", y = "No. identified proteins",
       color="Experiment")
ggplot(subset(dat.m, factor == "003_razor" | factor == "003_lfq")) +
  geom_boxplot(aes(x=variable, y=value, color=factor)) +
  labs(title="Comparison Rerun LFQ-Razor Experimt 003",
       x ="Extract", y = "No. identified proteins",
       color="Experiment")

boxplot(rsd_002_r,rsd_003_r,rsd_002,rsd_003,
        xlab="rerun experiment no.",
        names=c("002 razor","003 razor","002 lfq", "003 lfq"),
        ylab="relative sd")

```




**Results Razor**

```{r eval=FALSE, include=FALSE}
results_2gether_razor[c(1:4),-4]
```

**Results LFQ**

```{r eval=FALSE, include=FALSE}
results_2gether_lfq[c(1:4),-4]
```
